
## Regression Models Course Project
#### Benjamin Phillips

```{r}
library(dplyr)
library(ggplot2)
library(datasets)
```

```{r}
names(mtcars)
str(mtcars)
help(mtcars)
# [, 9]	 am	 Transmission (0 = automatic, 1 = manual)
```

From `help(mtcars)`, it's discovered that
 [, 9] is Transmission (0 = automatic, 1 = manual)


Make a new data frame for some exploratory analysis

```{r}
cars1 <- mtcars 
cars1$vs <- factor(cars1$vs, levels = c("0","1"), labels = c("automatic", "manual"))
```

```{r}
g1 <- ggplot(data = cars1,
             mapping = aes(x = vs, y = mpg)) + 
    geom_boxplot(fill = c("indianred","lightblue")) + 
    labs(title = "mpg for automatic and manual cars",
         x = "transmission")
g1

```

It appears from the plot that the mpg is very different for automatic and manual.
I have an feeling cylinder has a part to play as well.

```{r}
cars1$cyl <- factor(cars1$cyl, levels = c("4","6","8"))

g2 <- ggplot(data = cars1,
             mapping = aes(x = vs, y = mpg)) + 
    geom_boxplot(aes(fill = cyl)) + 
    labs(title = "mpg for automatic and manual cars, by cylinder size",
         x = "transmission")
g2
```

Now we're getting a different picture.

Let's try two models and compare them.

```{r}
mdl1 <- lm(formula = mpg ~ cyl, data = mtcars)
mdl1
summary(mdl1)

mdl2 <- lm(formula = mpg ~ cyl + vs, data = mtcars)
mdl2
summary(mdl2)
```

Adding vs (transmission) decreases the R-squared value indicating is doesn't improve our model.
Does adding vs (transmission) increase the variance?

```{r}
anova(mdl1, mdl2)
```

The deviance (RSS residual sim of squares) decreases slighty.

```{r}
g3 <- ggplot(data = mtcars,
             mapping = aes(x = cyl, y = mpg)) +
    geom_point(mapping = aes(colour = factor(vs, levels = c("0","1"), labels = c("automatic","manual")))) +
    geom_smooth(method = "lm") + 
    labs(title = "model 1", colour = "transmission")
g3
```


### Conclusion
The analysis suggests that transmission type is not a suitable variable to model mpg on. Instead, cylinder size is a better variable to use.

<br>
<br>
<br>
<br>
<br>


